---
title: "calciumSignalling_signatureClustering"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#block 1 
```

```{r}
#block 2 
# clear global enviroment                          
rm(list = ls())

#install.packages("devtools")


library(dtw)
library(dplyr)
library(ggplot2)
library(proxy)

#devtools::install_github(repo = "NMikolajewicz/scMiko",ref = "master",auth_token = "a3c1c9b15c496991c952d1fe3ccc52db770f22fa",force = TRUE)

library(scMiko)


```

```{r}
#Block 3 
#dir <- "D:/Users/Nick/Dropbox/PhD Projects - SVK/Papers/P2 receptor modeling - Anmar/R analysis/inputData/"
#file <- "input_master_v2.xlsx"

file <- "/Users/anthonyquint/Desktop/R analysis/inputData/input_master_v2.xlsx"

```


```{r import data, warning = F, message=F}
#Block 4 
#sheetNames <- openxlsx::getSheetNames(getLoadPath(file, dir))
sheetNames <- openxlsx::getSheetNames(file)

geneSets <- list()


for (i in 1:length(sheetNames)){
  current.sheet <-  sheetNames[i]
  gene.set <- readxl::read_excel(file, sheet = current.sheet)
  
  gene.set <- gene.set[complete.cases(gene.set), ]
  
  geneSets[[current.sheet]] <- gene.set 
}

```


```{r}

#Block 5 
# distMatrix <- dist(sample2, method=”DTW”)

plt.match <- F

if (plt.match){
  
  ts <-  geneSets[[5]]

time.vec <- as.vector((unlist(ts %>% dplyr::select(c("Time")))))

ts <- ts %>% dplyr::select(-c("Time"))

colnames(ts) <- paste("ROI", seq(1, ncol(ts)), sep = "")

i <- 1?
# j <- 3

# for (i in 1:ncol(ts)){
  for (j in 1:ncol(ts)){
    query <- as.vector(unlist(ts[, i]))
    reference <- as.vector(unlist(ts[, j]))
    alignment<-dtw(query,reference);

    # plot(alignment$index1,alignment$index2,main="Warping function");


    df.input <- data.frame(t = time.vec,
                           query = query,
                           ref = reference)

    loessMod10 <- loess(query ~ t, data=df.input, span=0.1) # 10% smoothing span
    #loessMod10 <- loess(query ~ t, data=df.input, span=0.03) # 5% smoothing span(Anthony's Edit)

    df.input$q.smooth <- predict(loessMod10)


    df.warped <- data.frame(t = time.vec[alignment$index1],
                            query = query[alignment$index1],
                            ref = reference[alignment$index2])

    plt.input <- df.input %>%
      ggplot() +
      geom_point(aes(x = t, y = query), color = "black") +
      geom_point(aes(x = t, y = reference), color = "red") +
      ggtitle(paste("Input (i=", i, ", j=", j,")",  sep = "")) +
      xlab("Time (s)") +
      ylab("Response") +
      theme_classic()



    plt.warped <- df.warped %>%
      ggplot() +
      geom_point(aes(x = t, y = query), color = "black") +
      geom_point(aes(x = t, y = ref), color = "red") +
      xlab("Time (s)") +
      ylab("Response") +
      theme_classic() +
      ggtitle("Output")

    print(cowplot::plot_grid(plt.input, plt.warped, ncol = 2))


    df.input %>%
      ggplot() +
      geom_line(aes(x = t, y = query), color = "black") +
      geom_line(aes(x = t, y = q.smooth), color = "red") +
      ggtitle(paste("Input (i=", i, ", j=", j,")",  sep = "")) +
      xlab("Time (s)") +
      ylab("Response") +
      theme_classic()
  }
# }

}








```


```{r}
#Block 6
ts.all <- NULL
for (i in 1:length(geneSets)){
  
  # get current condition
  cur.condition <- names(geneSets)[i]
  
  # get time series
  ts <-  geneSets[[i]]
  
  # smoothm time-series (noise reduction step)
  ts.smooth <- ts
  for (j in 2:ncol(ts)){
    cur.df <- data.frame(t = ts$Time, y = ts[ ,j])
    colnames(cur.df) <- c("t", "y")
        loessMod10 <- loess(y ~ t, data=cur.df, span=0.1) # 10% smoothing span
        #loessMod10 <- loess(y ~ t, data=cur.df, span=0.03) # 5% smoothing span(Anthony's Edit)
        
    ts.smooth[ ,j] <- predict(loessMod10) 
    
  }
  
  ts <- ts.smooth
  
  
  ts <- ts %>% dplyr::select(-c("Time"))
  
  # assign names to current timeseries
  colnames(ts) <- paste(cur.condition,".", seq(1, ncol(ts)), sep = "")
  
  if (!is.null(ts.all)){
    all.n <- nrow(ts.all)
    cur.n <- nrow(ts)
    
    if (cur.n > all.n){
      
      ts.na.filler <- ts.all[1:(cur.n-all.n), ]
      ts.na.filler[1:(cur.n-all.n), ] <- NA
      
      ts.all <- bind_rows(ts.all, ts.na.filler)
    } else  if (cur.n < all.n){
      ts.na.filler <- ts[1:(all.n-cur.n), ]
      ts.na.filler[1:(all.n-cur.n), ] <- NA
      ts <- bind_rows(ts, ts.na.filler)
    }
  }
  
  ts.all <- bind_cols(ts.all, ts)
  
}

ts.all <- ts.all[complete.cases(ts.all), ]


```


```{r calculate distance matrix}
#Block 7 
## Downloading all the extra things I need for WGCNA to work 
# if (!requireNamespace("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("GO.db")
# 
# if (!requireNamespace("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("impute")
# 
# if (!requireNamespace("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("preprocessCore")
## ^ Stuff I added in order to run WGCNA

library(WGCNA)


# get signautre matrix
resp.mat <- t(as.matrix(ts.all))
resp.norm <- BBmisc::normalize(resp.mat, method="standardize")


# calculate distance measures
distMatrix <- dist(resp.norm, method="DTW")

# labels for each signature
observedLabels <- colnames(ts.all)


```

```{r cluster data}

#Block 8 
# hierarchial clusters
hc <- hclust(distMatrix, method="average")
# hc <- hclust(distMatrix, method="complete")

# hc[["height"]] <- log(hc[["height"]])

# apply dynamic tree cut algorithm
# modules <- dynamicTreeCut::cutreeDynamicTree(dendro = hc, minModuleSize = 100, maxTreeHeight = 100, deepSplit = T)

modules <- dynamicTreeCut::cutreeDynamic(dendro = hc, distM = distMatrix, deepSplit = 4, pamRespectsDendro = F,
                        minClusterSize = 30, method = "tree")

# modules <- dynamicTreeCut::cutreeHybrid(dendro = hc, distM = distMatrix)

# assign colors to modules
module.colours = labels2colors(modules)

# plot dendrogram
plotDendroAndColors(hc, module.colours, 'Signatures', dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05, main='')

hist(log10(hc[["height"]]))

```


```{r}
#Block 9
module_colors= setdiff(unique(module.colours), "grey")
module.list <- list()
for (i in 1:length(module_colors) ){
  color <- module_colors[i]
  module.name <- paste("M", i, ".", color, sep = "")
  module.list[[module.name]]=observedLabels[which(module.colours==color)]
}

 table(module.colours)

```

```{r}
#Block 10 
# get module concentration scores
conc.legend <- c(-8, -7, -6, -5, -4, -3, -2)
names(conc.legend) <- c("e-8", "e-7", "e-6", "e-5", "e-4", "e-3", "e-2")


df.modScore <- data.frame(mod = names(module.list), score.mean = NA, score.median = NA, score.sd = NA)


for (i in 1:nrow(df.modScore)){
  a <- unlist(module.list[[i]])
  
  all.conc <- c()
  for (j in 1:length(conc.legend)){
    
    cur.conc <- rep(as.vector(conc.legend[j]),  sum(grepl(names(conc.legend)[j], a)) )
    
    all.conc <- c(all.conc, cur.conc)
  }
  
  
  
  df.modScore$score.mean[i] <- mean(all.conc)
  df.modScore$score.median[i] <- median(all.conc)
  df.modScore$score.sd[i] <- sd(all.conc)
  df.modScore$score.N[i] <- length(all.conc)
}

df.modScore <- df.modScore %>% arrange(score.mean)


```

```{r}

#Block 11 
df.modConc <- data.frame(module = names(unlist(module.list)), conc.label = as.vector(unlist(module.list)))

df.modConc$module <- as.character(df.modConc$module)

df.modConc$conc <- NA

for (j in 1:length(conc.legend)){
  df.modConc$conc[grepl(names(conc.legend)[j],  df.modConc$conc.label)] <- as.vector(conc.legend[j])
}

for (j in 1:length(module.list)){
  df.modConc$module[grepl(names(module.list)[j],  df.modConc$module)] <- names(module.list)[j]
}




df.modConc.sum <- df.modConc %>%
  dplyr::group_by(conc, module) %>%
  dplyr::tally()

df.modConc.sum <- df.modConc.sum %>%
  group_by(conc) %>%
  mutate(freq = n/sum(n))

df.modConc.sum$conc <- factor(df.modConc.sum$conc, levels = c(-8, -7, -6, -5, -4, -3, -2))
df.modConc.sum$module <- factor(df.modConc.sum$module, levels = as.character(df.modScore$mod))


df.modConc.sum$module 


df.modConc.sum %>% 
  ggplot(aes(x = conc, y = freq, fill = module)) + 
  geom_bar(stat="identity") + ylab("Module Frequency") + xlab("ATP concentration (logM)") + 
  ggtitle("Dose-Dependent Signature Frequency")


df.modConc.sum %>% 
  ggplot(aes(x = conc, y = freq, fill = module)) + 
  geom_bar(stat="identity") + 
  facet_wrap(~module) + 
  theme_classic()


```


```{r}
#Block 12 
# df.plot <- data.table(melt(data.table(class = as.factor(module.colours),
#                                         ts.all)))
# df.plot[, Time := rep(1:ncol(ts.all), each = nrow(ts.all))]
# df.plot[, ID := rep(1:nrow(ts.all), ncol(ts.all))]

library(reshape2)

ts.plot <- bind_cols(data.frame(time = 0.5*seq(1:nrow(ts.all))), ts.all)
ts.plot$time <-ts.plot$time - 0.5

ts.plot.long <- data.table::melt(ts.plot, id = "time", variable_name = "signature")


ts.plot.long$module <- NA
for (i in 1:length(module.list)){
 ts.plot.long$module[ts.plot.long$variable %in% module.list[[i]]] <-  names(module.list)[i]
}

 ts.plot.long$module[is.na( ts.plot.long$module)] <- "no.module"


```

```{r generate figure, fig.width= 10, fig.height=12}
#Block 13
# df.modScore <- df.modScore %>% arrange(score.mean)

ts.plot.long <- ts.plot.long[!(is.na(as.character(ts.plot.long$module))), ]

ts.plot.long$module <- factor(ts.plot.long$module, levels = as.character(df.modScore$mod))

ts.plot.long <- ts.plot.long[!(is.na(as.character(ts.plot.long$module))), ]

ts.plot.long$module <- factor(ts.plot.long$module, levels = as.character(df.modScore$mod))

# ts.plot.long.sub <-

# ts.plot.long$module[]
 
plt.sig <- ts.plot.long %>%
 ggplot(aes(x = time, y = value, group = variable)) +
  facet_wrap(~module, ncol = 4, scales = "fixed") +
  geom_line(color = "grey10", alpha = 0.1) +
  labs(x = "Time", y = "Load (normalised)") +
  theme_bw()

# plt.sig <- ts.plot.long %>%
#  ggplot(aes(x = time, y = value, group = variable)) +
#   facet_wrap(~module, ncol = 4, scales = "free_y") +
#   geom_line(color = "grey10", alpha = 0.1) +
#   labs(x = "Time", y = "Load (normalised)") +
#   theme_bw()

print(plt.sig)

```



```{r}

#Block 14 
# library(dtwclust)
# resp.norm <- BBmisc::normalize(resp.mat, method="standardize")
# 
# 
# resp.norm.sub <- resp.norm[sample(seq(1,nrow(resp.norm)), 100), ]
# 
# 
# cluster.n <- 2L:17L
# clust.pam <- dtwclust::tsclust(resp.norm.sub, type="partitional", k=cluster.n, distance="dtw", centroid="pam")
# 
# # dtwclust::cvi(clust.pam[[1]])
# res <- lapply(clust.pam,cvi )
# 
# 
# df.res <- NULL
# for (i in 1:length(res)){
#   cur.res <- data.frame(Sil = res[[i]][1],
#              SF = res[[i]][2],
#              CH = res[[i]][3],
#              DB = res[[i]][4],
#              DBstar = res[[i]][5],
#              D = res[[i]][6])
#   
#   df.res <- bind_rows(df.res, cur.res)
# }
# 
# df.res <- bind_cols(data.frame(n.clust = cluster.n), df.res)
# 
# plt.sil <- df.res %>% ggplot(aes(x = n.clust, y = Sil)) + geom_point() + ggtitle("Sil")
# plt.sf <- df.res %>% ggplot(aes(x = n.clust, y = SF)) + geom_point() + ggtitle("SF")
# plt.ch <- df.res %>% ggplot(aes(x = n.clust, y = CH)) + geom_point() + ggtitle("CH")
# plt.db <- df.res %>% ggplot(aes(x = n.clust, y = DB)) + geom_point() + ggtitle("DB")
# plt.dbstar <- df.res %>% ggplot(aes(x = n.clust, y = DBstar)) + geom_point() + ggtitle("DBstar")
# plt.d <- df.res %>% ggplot(aes(x = n.clust, y = D)) + geom_point() + ggtitle("D")
# 
# cowplot::plot_grid(plt.sil, plt.sf, plt.ch, plt.db, plt.dbstar, plt.d, ncol = 3)
# 
# clust.pam.v2 <- tsclust(resp.norm.sub, type="partitional", k=8L, distance="dtw", clustering="pam")
# 
# 
# # clust.hier <- dtwclust::tsclust(resp.norm.sub, type = "h", k = 8L, distance = "dtw")
# # 
# # 
# # 
# plot(clust.pam.v2)
# 
# 
# plot(clust.pam.v2, type="sc")
# 
# plot(clust.pam.v2, type = "sc", clus =8)

# table(cutree(clust.hier, k=8L))




```



```{r}
#Chunk 15
# Gonna try to get all the clustering data in one spreadsheet 

#plt.sig[[1]]

#View(plt.sig[["data"]])

#(plt.sig[["data"]])[,3]

#install.packages("xlsx")
library(openxlsx)

arrange(plt.sig[["data"]], module)

openxlsx::write.xlsx(arrange(plt.sig[["data"]], module), "/Users/anthonyquint/Desktop/blah.xlsx", sheetName = "Data",col.names = TRUE, row.names = TRUE, append = FALSE)


#openxlsx::write.xlsx(plt.sig[["data"]], "/Users/anthonyquint/Desktop/Clusters_in_Excel.xlsx", sheetName = "Data",col.names = TRUE, row.names = TRUE, append = FALSE)


# "/Users/anthonyquint/Desktop/Clusters_in_Excel.xlsx"




